The influence of entry regulation on labour market formalization

### 2023-03-25 proposal/proposal.qmd
Finish introduction.

This political instability has also made the deregulation of the labour market difficult to discuss in the legislative branch: only one set of market-friendyl policies set forth by the new president have been passed, the labour market reform being immediately rejected by a strong majority in Congress. Given the difficulty to pass any kind of labour-related policy, other types of reforms which indirectly target the labour market would be welcome. The present analysis attempts to evaluate a policy like this one, which was not directly aimed to enhance employment, but still could cause an effect if companies with informal employees decide to formalize job contracts after entering the formal sector of the economy. The policy reform allowed a new type of legal business to exist, the *Sociedad de Accciones Simplificada* (SAS), which also exists in several other countries in the region. This type of business is known to be much less burdensome than other types of companies like anonymous societies or limited liability companies. Specifically, a SAS business type can be created with considerably lower monetary and time costs (notarization, registering, etc.) as well as with lower amounts of capital contributions. It is sensible to believe that the reform was unexpected and exogeneous since, while the law that allowed for this reform came into force in early 2020, the law did not specifically plan for the creation of this type of business, only allowed for its creation under the Ecuadorian corporate law [@CaminoMogro.2022].

@@

The literature on informality is extensive. The theoretical framework points to informality being an endogenous decision, where firms with low productivity often prefer to remain informal, the size of the firm often being an important determinant of this decision [@Delechat2020; @MauricioPrado.2011; @ElianeElBadaoui.2010]. The fact that informality is an endogenous decision makes informality difficult to define in a specific setting; many informal workers are not officially registered and may not report their activities accurately [Oviedo.2009].

>>>>

This political instability has also made the deregulation of the labour market difficult to discuss in the legislative branch: only one set of market-friendyl policies set forth by the new president have been passed, the labour market reform being immediately rejected by a strong majority in Congress. Given the difficulty to pass any kind of labour-related policy, other types of reforms which indirectly target the labour market would be welcome. The present analysis attempts to evaluate a policy like this one, which was not directly aimed to enhance employment, but still could cause an effect if companies with informal employees decide to formalize job contracts after entering the formal sector of the economy.

The policy reform allowed a new type of legal business to exist, the *Sociedad de Accciones Simplificada* (SAS), which also exists in several other countries in the region. This type of business is known to be much less burdensome than other types of companies like anonymous societies or limited liability companies. Specifically, a SAS business type can be created with considerably lower monetary and time costs (notarization, registering, etc.) as well as with lower amounts of capital contributions. It is sensible to believe that the reform was unexpected and exogeneous since, while the law that allowed for this reform came into force in early 2020, the law did not specifically plan for the creation of this type of business, only allowed for its creation under the Ecuadorian corporate law [@CaminoMogro.2022].

@@

The literature on informality is extensive. The theoretical framework points to informality being an endogenous decision, where firms with low productivity often prefer to remain informal, the size of the firm often being an important determinant of this decision [@Delechat2020; @MauricioPrado.2011; @ElianeElBadaoui.2010]. The fact that informality is an endogenous decision makes informality difficult to investigate empirically; many informal workers are not officially registered and may not report their activities accurately [Oviedo.2009].
